---
title: "HRP_ETH Moire"
author: "William Louie"
date: "`r Sys.Date()`"
output: html_document
---

### Load libraries and working directory

```{r load libraries, echo=FALSE, warning = FALSE, message= FALSE}

library(moire)
library(dplyr)
library(magrittr)
library(ggplot2)
library(reshape2)
library(tidyr)
library(tidyverse)
library(stringr)
library(tools)
library(purrr)
library(parallel)
library(data.table)
library(vctrs)
library(plotly)
library(gridExtra)
library(scales)
library(gtsummary)
library(gt)
library(ggpubr)
library(dcifer)
library(patchwork)

```

### Import moire results and metadata

```{r import, echo=FALSE, warning = FALSE, message= FALSE}

setwd("/Users/williamlouie/Library/CloudStorage/Box-Box/HRP_Ethiopia/HRP_analysis")

# metadata with qpcr
metadata_qpcr <- read_csv("/Users/williamlouie/Library/CloudStorage/Box-Box/HRP_Ethiopia/metadata/HRP_metadata.csv") %>%
  mutate(sample_id = as.character(QR_code)) %>%
  select(sample_id, pfqparul, pvqcopyul)
# New metadata with reclassifications
metadata <- read_csv("/Users/williamlouie/Library/CloudStorage/Box-Box/HRP_Ethiopia/metadata/EthiopiaHRP_qPCRpos_collated_26Jan2025.csv") %>%
  mutate(sample_id = as.character(participantID)) %>%
  select(sample_id, PopClass, migrantagworker, SeasSAWorker)

# Input
hrp_moire_input <- read.csv("./hrp_moire/FINAL_moire_input_012725.csv")
# dat <- moire::load_long_form_data(hrp_moire_input)

# If spitting by site
hrp_moire_input_hl <- hrp_moire_input %>%
  filter(PopClass == "Highlands")
dat_hl <- moire::load_long_form_data(hrp_moire_input_hl)

hrp_moire_input_ll <- hrp_moire_input %>%
  filter(PopClass == "Lowlands")
dat_ll <- moire::load_long_form_data(hrp_moire_input_ll)

# Import moire results
moire_results_hl <- readRDS("./hrp_moire/NEW_moire_results_hl.rds")
moire_results_ll <- readRDS("./hrp_moire/NEW_moire_results_ll.rds")


```

### Combine & summarize
Fig 2a: COI/eCOI distribution by population

```{r summarize, echo=TRUE}

# COI summaries
## summarize COI for HL and LL separately too
coi_hl <- moire::summarize_coi(moire_results_hl)
ecoi_hl <- moire::summarize_effective_coi(moire_results_hl)
coi_summary_hl <- coi_hl %>%
  full_join(ecoi_hl, by = "sample_id")
coi_ll <- moire::summarize_coi(moire_results_ll)
ecoi_ll <- moire::summarize_effective_coi(moire_results_ll)
coi_summary_ll <- coi_ll %>%
  full_join(ecoi_ll, by = "sample_id")
coi_summary_comb <- rbind(coi_summary_hl, coi_summary_ll)
coi_summary_comb <- coi_summary_comb %>%
  mutate(sample_id = as.character(sample_id)) %>%
  left_join(metadata, by = "sample_id") %>%
  left_join(metadata_qpcr, by = "sample_id")

# Get p-values
variables_to_test <- c("post_coi_mean", "post_coi_med", "post_effective_coi_mean", "post_effective_coi_med")
# Function to perform t-test on all variables listed
perform_t_tests <- function(var) {
  coi_summary_comb %>%
    select(PopClass, !!sym(var)) %>%
    drop_na(!!sym(var)) %>%  # Remove NAs
    expand(PopClass1 = PopClass, PopClass2 = PopClass) %>%
    filter(PopClass1 < PopClass2) %>%
    rowwise() %>%
    mutate(
      t_test = list(t.test(
        formula = as.formula(paste(var, "~ PopClass")),  # Dynamic formula
        data = coi_summary_comb %>% filter(PopClass %in% c(PopClass1, PopClass2))
      )),
      p_value = t_test$p.value,
      t_value = t_test$statistic
    ) %>%
    ungroup() %>%
    select(PopClass1, PopClass2, p_value, t_value) %>%  # Keep necessary columns
    mutate(
      variable = var,  # Store variable name
      p_value = signif(p_value, 3),  # Round p-values
      t_value = round(t_value, 3),
      significance = case_when(
        p_value < 0.001 ~ "***",
        p_value < 0.01  ~ "**",
        p_value < 0.05  ~ "*",
        TRUE            ~ "ns"))
}
# Run t-tests for all variables and combine results
t_test_results <- bind_rows(lapply(variables_to_test, perform_t_tests))


write_csv(coi_summary_comb, "NEW_coi_summary.csv")
write_csv(t_test_results, "STATS_coi_fig2a.csv")


coi_summary_stats <- coi_summary_comb %>%
  group_by(PopClass) %>%
  summarise(
    total_samples = n(),
    mean_coi_mean = mean(post_coi_mean),
      sd_coi_mean = sd(post_coi_mean),
      se_coi_mean = sd(post_coi_mean) / sqrt(total_samples),
      lower_ci_coi_mean = mean_coi_mean - 1.96 * se_coi_mean,
      upper_ci_coi_mean = mean_coi_mean + 1.96 * se_coi_mean,
    mean_coi_med = mean(post_coi_med),
      sd_coi_med = sd(post_coi_med),
      se_coi_med = sd(post_coi_med) / sqrt(total_samples),
      lower_ci_coi_med = mean_coi_med - 1.96 * se_coi_med,
      upper_ci_coi_med = mean_coi_med + 1.96 * se_coi_med,
    `Poly count (COI)` = sum(post_coi_med > 1),
    `Percent poly (COI)` = (`Poly count (COI)` / total_samples) * 100,
    mean_ecoi_mean = mean(post_effective_coi_mean),
      sd_ecoi_mean = sd(post_effective_coi_mean),
      se_ecoi_mean = sd(post_effective_coi_mean) / sqrt(total_samples),
      lower_ci_ecoi_mean = mean_ecoi_mean - 1.96 * se_ecoi_mean,
      upper_ci_ecoi_mean = mean_ecoi_mean + 1.96 * se_ecoi_mean,
    mean_ecoi_med = mean(post_effective_coi_med),
      sd_ecoi_med = sd(post_effective_coi_med),
      se_ecoi_med = sd(post_effective_coi_med) / sqrt(total_samples),
      lower_ci_ecoi_med = mean_ecoi_med - 1.96 * se_ecoi_med,
      upper_ci_ecoi_med = mean_ecoi_med + 1.96 * se_ecoi_med,
    `Poly count (eCOI)` = sum(post_effective_coi_mean > 1.1),
    `Percent poly (eCOI)` = (`Poly count (eCOI)` / total_samples) * 100,
    .groups = 'drop'
  )

write.csv(coi_summary_stats, "STATS_coi_summary_all.csv")



# Population colors
pop_color <- c("Highlands" = "darkgreen", "Lowlands" = "purple")

# COI mean summary
fig2a <- ggplot(coi_summary_comb, aes(x = post_coi_mean)) + 
  geom_histogram(
    binwidth = 1, 
    aes(fill = PopClass)) +
  scale_x_continuous(breaks = seq(1, 9, 1)) +
  geom_histogram( # Add eCOI histogram
    aes(x = post_effective_coi_mean),
    binwidth = 0.25,
    color = "grey", fill = "grey",       # Set color to grey
    alpha = 0.5) +
    geom_vline(data = coi_summary_stats, aes(xintercept = mean_coi_med), color = "black", linetype = 'dashed', size = 0.5) + # Add mean COI line
    geom_vline(data = coi_summary_stats, aes(xintercept = mean_ecoi_med), color = "black", linetype = 'dotted', size = 0.5) +
  scale_fill_manual(values = pop_color, guide = "none") +
  xlab("Mean COI/ eCOI") +
  ylab("Count") +
  theme_classic() +
  facet_wrap(~PopClass, scales = "free", nrow = 2)
fig2a
ggsave("Plot_coi_distribution.jpeg", plot = fig2a, width = 4, height = 6, dpi = 300)


```

### Species correlations

```{r species, echo=FALSE, warning = FALSE, message= FALSE} 

species_loci <- read_csv("/Users/williamlouie/Library/CloudStorage/Box-Box/HRP_Ethiopia/HRP_moire_setup/FINAL_all_loci_012125.csv") %>%
  filter(Category == "Species") %>%
  group_by(sample_id, Locus) %>%
  summarise(total_reads = sum(Reads))

pv <- species_loci %>%
  filter(grepl("Pv", Locus))
pv_samples <- pv$sample_id

# Calculate Pf to Pv ratio
ratioPfPv <- species_loci %>%
        filter(grepl("Pv|Pf", Locus)) %>%
        pivot_wider(names_from = Locus, values_from = total_reads) %>%
        mutate(across(everything(), ~replace_na(.x, 0))) %>%
        mutate(reads_ratio = `PvP01_12_v1-1184983-1185208-1AB`/`Pf3D7_13_v3-1041593-1041860-1AB`)

```

### Individual coi, effective coi, within-host relatedness
for troubleshooting, combining data

```{r moire COI_eCOI, echo=FALSE, warning = FALSE, message= FALSE} 

# Plot individual coi median
g1 <- ggplot(coi_summary_comb, aes(x = as.character(sample_id), y = post_coi_med,
                           ymin = post_coi_lower,
                           ymax = post_coi_upper)) +
    geom_errorbar() +
    geom_point() +
    theme_classic(base_size = 10) +
    theme(axis.text.x = element_blank()) +
    facet_wrap(~PopClass, scales = "free") +
    xlab("Sample code")
g1

# Plot individual coi mean
g2 <- ggplot(coi_summary_comb, aes(x = as.character(sample_id), y = post_coi_mean,
                           ymin = post_coi_lower,
                           ymax = post_coi_upper)) +
    geom_errorbar() +
    geom_point() +
    theme_classic(base_size = 10) +
    theme(axis.text.x = element_blank()) +
    facet_wrap(~PopClass, scales = "free") +
    xlab("Sample code")
g2


# Plot individual effective coi median
g3 <- ggplot(coi_summary_comb, aes(x = as.character(sample_id), y = post_effective_coi_med,
                           ymin = post_effective_coi_lower,
                           ymax = post_effective_coi_upper)) +
    geom_errorbar() +
    geom_point() +
    theme_classic(base_size = 10) +
    theme(axis.text.x = element_blank()) +
    facet_wrap(~PopClass, scales = "free") +
    xlab("Sample code")
g3

# Plot individual effective coi mean
g4 <- ggplot(coi_summary_comb, aes(x = as.character(sample_id), y = post_effective_coi_mean,
                           ymin = post_effective_coi_lower,
                           ymax = post_effective_coi_upper)) +
    geom_errorbar() +
    geom_point() +
    theme_classic(base_size = 10) +
    theme(axis.text.x = element_blank()) +
    facet_wrap(~PopClass, scales = "free") +
    xlab("Sample code")
g4

```

### COI distribution
Fig 2b: correlation between COI and eCOI
Fig 2c: species distribution
Combines panels to make final Figure 2

```{r coi_deeper, echo=FALSE, warning = FALSE, message= FALSE}

# Add species info
coi_summary_comb %<>%
  mutate(pv_status = case_when(
    sample_id %in% pv_samples & pvqcopyul > 0 ~ "Pf + Pv",
    .default = "Pf"
  )) %>%
  mutate(clonality = case_when(
    post_coi_med > 1 ~ "polyclonal",
    post_coi_med == 1 ~ "monoclonal"
  ))

  
# Count for N to annotate
pop_counts <- coi_summary_comb %>%
  count(PopClass, pv_status, clonality) %>%
  group_by(PopClass) %>%
  mutate(percentage = round(n / sum(n) * 100, 1))
# Plot
color_clone <- c("monoclonal" = "skyblue", "polyclonal" = "brown")
fig2c <- ggplot(coi_summary_comb, aes(x = PopClass, y = pfqparul)) +
  geom_boxplot(
    aes(fill = clonality), 
    alpha = 0.3, 
    outlier.shape = NA, 
    position = position_dodge(width = 0.6)) +
  geom_jitter(
    aes(color = clonality),
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.6),
    size = 1, alpha = 0.5) +
  scale_fill_manual(values = color_clone) +
  scale_color_manual(values = color_clone) +
  scale_y_log10(labels = scales::trans_format("log10", math_format(10^.x))) +
  facet_wrap(~pv_status) +
  theme_bw() +
  theme(
    legend.position = "right",
    axis.title.x = element_blank()) +
  ylab(expression("Pf p/μL"))

fig2c
ggsave(plot = fig2c, "Plot_clonality_parasitemia.jpeg", height = 8, width = 6, dpi = 300)



# Plot correlation between coi and eff_coi, by population
max_coi <- round(max(coi_summary_comb$post_coi_mean))

ggplot(coi_summary_comb, aes(x = post_coi_mean, y = post_effective_coi_mean, color = PopClass)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  geom_abline(slope = 1, intercept = 0, linetype = "solid", color = "gray") +
  facet_wrap(~ PopClass) +
  scale_color_manual(values = c("Highlands" = "darkgreen", "Lowlands" = "purple")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, max_coi + 1), breaks = seq(1, max_coi + 1, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, max_coi + 1), breaks = seq(1, max_coi + 1, 1)) +
  xlab("COI") + 
  ylab("eCOI") +
  theme_classic()

# Test whether the linear regression is different
# Fit linear models by population
library(broom)
# Tidy linear model output by population
lm_by_pop <- coi_summary_comb %>%
  group_by(PopClass) %>%
  do(tidy(lm(post_effective_coi_mean ~ post_coi_mean, data = .)))
# View model coefficients (intercept and slope per population)
lm_by_pop
lm_slopes <- lm_by_pop %>%
  select(PopClass, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename(intercept = `(Intercept)`, slope = `post_coi_mean`)
# 1) Extract slope/intercept as above
# 2) Create equation labels as character strings
lm_slopes <- lm_slopes %>%
  mutate(
    equation_label = paste0("y = ", round(intercept, 2), " + ", round(slope, 2), "x"))
# 3) Merge with PopClass so we can color the text accordingly
label_data <- coi_summary_comb %>%
  distinct(PopClass) %>%
  left_join(lm_slopes, by = "PopClass") %>%
  mutate(label_color = ifelse(PopClass == "Highlands", "darkgreen", "purple")) %>%
  mutate(label_x = 0.5,  # left side of plot
         label_y = seq(from = max_coi + 0.8, by = -0.7, length.out = n())) 
# New plot
fig2b <- ggplot(coi_summary_comb, aes(x = post_coi_mean, y = post_effective_coi_mean, color = PopClass)) +
  geom_point() +
  geom_smooth(aes(group = PopClass), method = "lm", se = FALSE, color = "black", linetype = "solid") +
  # geom_text(
  #   data = label_data,
  #   aes(x = label_x, y = label_y, label = equation_label),
  #   color = label_data$label_color,
  #   hjust = 0,
  #   inherit.aes = FALSE
  # ) +
  scale_color_manual(values = c("Highlands" = "darkgreen", "Lowlands" = "purple")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, max_coi + 1), breaks = seq(1, max_coi + 1, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, max_coi + 1), breaks = seq(1, max_coi + 1, 1)) +
  xlab("COI") +
  ylab("eCOI") +
  theme_classic()
fig2b
ggsave(plot = fig2b, "Plot_coi_ecoi_correl.jpeg", height = 6, width = 7, dpi = 300)

## STATS
# Is the slope significantly different?
lm_interaction <- lm(post_effective_coi_mean ~ post_coi_mean * PopClass, data = coi_summary_comb)
summary(lm_interaction) # Not statistically different slope
# Save this stat
lm_summary <- summary(lm_interaction)$coefficients
lm_summary_df <- as.data.frame(lm_summary)
write.csv(lm_summary_df, "STAT_coi_ecoi_interaction_pop.csv", row.names = TRUE)


# Combine Figure 2 panels
fig2bc <- ggarrange(fig2b, fig2c, nrow = 2, labels = c("b", "c"))
fig2 <- ggarrange(fig2a, fig2bc, ncol = 2, widths = c(3, 4), labels = c("a", ""))
# Save Figure 2
ggsave(plot = fig2, "Fig2_coi_general.jpeg", height = 6, width = 9, dpi = 300)


# Here we make the gtsummary table 
coi_arrange <- coi_summary_comb %>%
  mutate(`Poly count (COI)` = post_coi_med > 1) %>%
  mutate(`Poly count (eCOI)` = post_effective_coi_med > 1.1) %>%
  rename(`COI mean` = post_coi_mean) %>%
  rename(`COI median` = post_coi_med) %>%
  rename(`eCOI mean` = post_effective_coi_mean) %>%
  rename(`eCOI median` = post_effective_coi_med) %>%
  rename(`naive COI` = naive_coi) %>%
  mutate(`naive COI` = as.numeric(`naive COI`))
# Create summary table for COI
table_coi <- coi_arrange %>%
  tbl_summary(
    by = PopClass,
    include = c(`COI mean`, `COI median`, `eCOI mean`, `eCOI median`, 
                `Poly count (COI)`, `Poly count (eCOI)`),
    statistic = list(
      `COI mean` ~ "{mean} ({sd})",
      `COI median` ~ "{mean} ({sd})",
      `eCOI mean` ~ "{mean} ({sd})",
      `eCOI median` ~ "{mean} ({sd})",
      `Poly count (COI)` ~ "{n} ({p}%)",
      `Poly count (eCOI)` ~ "{n} ({p}%)"
    )
  ) %>%
  add_p() %>%
  modify_header(label = "posterior COI")
table_coi

b <- table_coi %>%
  as_gt()
# Save the gt table as an image
gtsave(b, filename = "Table_coi_distribution.pdf")


```

### Heterozygosity
compare median He between HL and LL

```{r heterozygosity, echo=FALSE, warning = FALSE, message= FALSE}

# Get He from moire results
## HL
moire_results_hl$chains[[1]]$allele_freqs <- lapply(moire_results_hl$chains[[1]]$allele_freqs, function(af_chain) {
     lapply(af_chain, function(sample) {
         replace(sample, which(is.na(sample)), 0)
     })
 })
he_hl <- summarize_he(moire_results_hl)
he_hl$PopClass <- "Highlands"
hist(he_hl$post_stat_mean)

## LL
moire_results_ll$chains[[1]]$allele_freqs <- lapply(moire_results_ll$chains[[1]]$allele_freqs, function(af_chain) {
     lapply(af_chain, function(sample) {
         replace(sample, which(is.na(sample)), 0)
     })
 })
he_ll <- summarize_he(moire_results_ll)
he_ll$PopClass <- "Lowlands"
hist(he_ll$post_stat_mean)

# Merge Hl and LL
he_merge <- rbind(he_hl, he_ll)



# Do summary stats
he_merge_summary <- he_merge %>%
  group_by(PopClass) %>%
  summarise(total_n = n(),
    mean_he = mean(post_stat_mean),
      sd_mean = sd(post_stat_mean),
      se_mean = sd(post_stat_mean) / sqrt(total_n),
      lower_ci_mean = mean_he - 1.96 * se_mean,
      upper_ci_mean = mean_he + 1.96 * se_mean,
    median_he = median(post_stat_med),
      sd_med = sd(post_stat_med),
      se_med = sd(post_stat_med) / sqrt(total_n),
      lower_ci_med = median_he - 1.96 * se_med,
      upper_ci_med = median_he + 1.96 * se_med)

# Wilcoxon test to see if median He between HL and LL are different
# Ensure PopClass is a factor and remove NAs
he_clean <- he_merge %>%
  filter(!is.na(post_stat_med) & !is.na(PopClass)) %>%
  mutate(PopClass = as.factor(PopClass))
wilcox_pvalue <- wilcox.test(post_stat_med ~ PopClass, data = he_clean)$p.value
he_stats <- tibble(
  wilcox_pvalue = wilcox_pvalue)

write_csv(he_merge, "heterozygosity/heterozygosity_all_loci.csv")
write_csv(he_merge_summary, "heterozygosity/heterozygosity_summary.csv")


```

### Allele frequency (done on Arwen)
Here we calculate the naive allele frequencies of HL and LL separately, then use a weighted average. A single input for allele frequencies is currently needed to feed into Dcifer. We will call this afreq and this will be used to run Dcifer in the subsequent chunk. I ran this on Arwen already, so can just import (NEW_hrp_allele_freq_avg.csv)

```{r allele freq, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# Import
afreq <- read_csv("/Users/williamlouie/Library/CloudStorage/Box-Box/HRP_Ethiopia//HRP_analysis/hrp_moire/NEW_hrp_allele_freq_avg.csv")


# Naive AF
naive_AF_list_hl <- calculate_naive_allele_frequencies(dat_hl$data)
naive_AF_table_hl <- naive_AF_list_hl %>%
  map_df(as_tibble) %>%
  mutate(target = map_depth(naive_AF_list_hl, 1, names) %>% unlist(use.names = F)) %>%
  rename(naive_AF.hl = value) %>%
  mutate(locus = str_extract(target, "^[^.]+")) %>%
  rename(allele = target)
# Moire AF
AF_list_hl <- calculate_median_afs(moire_results_hl)
AF_table_hl <- AF_list_hl %>%
  map_df(as_tibble) %>%
  mutate(target = map_depth(naive_AF_list_hl, 1, names) %>% unlist(use.names = F)) %>%
  rename(mean_AF.hl = value) %>%
  mutate(locus = str_extract(target, "^[^.]+")) %>%
  rename(allele = target) %>%
  select(-locus) %>%
  full_join(naive_AF_table_hl, by = "allele")

# Now do the same for LL
# Naive AF
naive_AF_list_ll <- calculate_naive_allele_frequencies(dat_ll$data)
naive_AF_table_ll <- naive_AF_list_ll %>%
  map_df(as_tibble) %>%
  mutate(target = map_depth(naive_AF_list_ll, 1, names) %>% unlist(use.names = F)) %>%
  rename(naive_AF.ll = value) %>%
  mutate(locus = str_extract(target, "^[^.]+")) %>%
  rename(allele = target)
# Moire AF
AF_list_ll <- calculate_median_afs(moire_results_ll)
AF_table_ll <- AF_list_ll %>%
  map_df(as_tibble) %>%
  mutate(target = map_depth(naive_AF_list_ll, 1, names) %>% unlist(use.names = F)) %>%
  rename(mean_AF.ll = value) %>%
  mutate(locus = str_extract(target, "^[^.]+")) %>%
  rename(allele = target) %>%
  select(-locus) %>%
  full_join(naive_AF_table_ll, by = "allele")

# Merge together HL and LL
# Average moire AF
AF_merge <- AF_table_hl %>%
  full_join(AF_table_ll, by = "allele") %>%
  mutate(mean_AF.hl = ifelse(is.na(mean_AF.hl), 0, mean_AF.hl),
         mean_AF.ll = ifelse(is.na(mean_AF.ll), 0, mean_AF.ll)) %>%
  mutate(avg_AF = ((mean_AF.hl + mean_AF.ll) / 2)) %>%
  rowwise() %>%
  mutate(sd = sd(c(mean_AF.hl, mean_AF.ll), na.rm = TRUE)) %>%
  mutate(locus = coalesce(locus.x, locus.y)) %>%
  select(-c(locus.x, locus.y))
write.csv(AF_merge, "hrp_allele_freq_avg.csv")

# Format allele frequencies (avg) for Dcifer
afreq <- formatAfreq(AF_merge, lvar = "locus", avar = "allele", fvar = "avg_AF")
write_rds(afreq, "./hrp_dcifer/mean_allele_frequencies.rds")

# Plot AF
plot_AF <- function(AF_merge){
  g <- ggplot(AF_merge) +
    geom_point(aes(y = mean_AF.hl, x = forcats::fct_reorder(allele, avg_AF), color = "Highlands"), alpha = .2, show.legend = TRUE) +
    geom_point(aes(y = mean_AF.ll, x = forcats::fct_reorder(allele, avg_AF), color = "Lowlands"), alpha = .2, show.legend = TRUE) +
    geom_point(aes(y = avg_AF, x = forcats::fct_reorder(allele, avg_AF), color = "Average"), size = 1, alpha = 1, show.legend = TRUE) +
    ylab("Allele Frequency") +
    xlab("Allele") +
    theme_classic(base_size = 12) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1, size = 5.5)) +
    expand_limits(x = 0, y = 0)
    ggtitle("Allele Frequency") +
    scale_color_manual(values = c("Highlands" = "blue", "Lowlands" = "red", "Average" = "black"))
  g
}

plot_AF(AF_merge)



```



